---
title: "vaers and covid"
author: "Dan Weinberger"
date: "3/1/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(data.table)
```

#We will read in the data from VAERS For Jan-July 2021 (accessd August 2, 2021)

```{r cars}
#vaers2021 <- read.csv('./Data/2021VAERSDATA.csv')
# vaers2021 <- read.csv('./Data/2021VAERSDATA.csv')
# vaers2021vax <- read.csv('./Data/2021VAERSVAX.csv')
# vaers2021symp <- read.csv('./Data/2021VAERSSYMPTOMS.csv')
# comb1 <- merge(vaers2021, vaers2021vax, by='VAERS_ID')
# comb2 <- merge(comb1, vaers2021symp, by='VAERS_ID')
# covid <- comb2[comb2$VAX_TYPE=='COVID19',]
# 
# saveRDS(covid,'./Data/2021vaers_covid.rds')

#write.csv(mmr2019,'./Data/2019V_mmr_vaers.csv')
a1<-readRDS('./Data/2021vaers_covid.rds')

```

Print first 10 symptom complaints
```{r}
a1$SYMPTOM_TEXT[1:10]
```


## See summaries of categories
```{r}
  sort(                  #sort 
    table(a1$SYMPTOM1),#create summary tabkle  
    decreasing=T)[1:100]  #order in decreasing order
```

## Remove duplicate records

Take the first entry for each VAERS_ID

Sort the data by ID
```{r}
a1 <- a1[order(a1$VAERS_ID),]
```

Convert to a 'data.table' format, which is computationally much faster in R
```{r}
a2 <- data.table(a1, key = "VAERS_ID")
```

Keep just the first observation using head(1)
```{r}
a2 <- a2[, head(.SD, 1), by = key(a2)]
```

Check number of rows in original and new dataset
```{r}
nrow(a1)
nrow(a2)
```

## Creating a case definition



## Let's flag all people potentially with myocarditis/pericarditis
the grep() function searches text strings.
toupper makes all of the letters CAPITAL to avoid issues with upper/lower case when matching
```{r}
a2$mycocarditis<-0  #initializae the variable called "rash"

a2$mycocarditis[
  grep('MYOCARDITIS',   #word you want to match
       toupper(a2$SYMPTOM_TEXT)) #search the symptom text for the keyword
      ]   <-1   #flag it as 1

a2$pericardiditis <- 0

a2$pericardiditis[
  grep('PERICARDITIS',   #word you want to match
       toupper(a2$SYMPTOM_TEXT)) #search the symptom text for the keyword
      ]   <-1   #flag it as 1
```

How many people have 'pericarditis' or 'myocarditis' listed in the free text?

```{r}
table(a2$mycocarditis, a2$pericardiditis)

```

View the records for people with this flag

```{r}
View(a2[a2$mycocarditis==1 | a2$pericardiditis==1,])


a3 <- a2[a2$mycocarditis==1 | a2$pericardiditis==1,]
```






## see age distribution of reports
```{r}
par(mfrow=c(1,2))
hist(a2$AGE_YRS, xlim=c(0,120), main='Age of all complaintants')

hist(a3$AGE_YRS, xlim=c(0,120), main='Myocarditis/pericarditis age')

```

```{r}
hist(a3$NUMDAYS[!is.na(a3$NUMDAYS) & a3$NUMDAYS<60], xlim=c(0,60), breaks=10, main='Time of symptom onset')
```


